# Highly Migratory Species Stock Status {#hms_stock_status}

**Description**: Summary of the most recent stock assessment results for each assessed HMS species.

**Indicator family**: 

- [X] Megafauna


**Contributor(s)**: Jennifer Cudney

**Affiliations**: NEFSC

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator
This section summarizes Highly Migratory Species Stock Status as it is presented in the latest Atlantic Highly Migratory Species Stock Assessment and Fishery Evaluation (SAFE) report. Stock assessments for Atlantic HMS are completed through one of three processes. Most Atlantic HMS stocks that are managed internationally under the authority of the MSA and ATCA (tunas, billfish, and swordfish) are assessed by the ICCAT SCRS. Given their migratory nature and because of their interactions with ICCAT fisheries, some pelagic sharks are also assessed by the ICCAT SCRS. Other shark stocks that do not interact with ICCAT fisheries are assessed via the SouthEast Data Assessment and Review (SEDAR) process. SEDAR assessments are completed for domestically managed HMS shark stocks. Sharks may also be assessed through an independent stock assessment is completed by a third party and approved for use in management by NOAA Fisheries (“external assessments”). Unless otherwise indicated, most assessments presented herein are done through either the ICCAT or SEDAR process.

## Key Results and Visualizations
The domestic and international status of 28 species or stocks are described in the most recent HMS SAFE report. International and domestic HMS stock status may not be identical, based on different thresholds and status determination criteria used at the domestic and international levels (see Figures). 

Regarding international stock status, overfishing is either not likely to be occurring or not occurring for 3 and 7 species/stocks, respectively. Overfishing is occurring for 3 species/stocks, and is undetermined for 1 stock. Regarding domestic stock status, overfishing is not occurring for 18 species/stocks. Overfishing is occurring for 5 species/stocks, and is unknown for 4 species/stocks

Regarding international stock status, 2 species/stocks are likely not overfished, 4 species/stocks are not overfished, and 6 species/stocks are considered overfished. Regarding domestic stock status, 13 species/stocks are not considered overfished (2 of these are either rebuilding or are rebuilt), 9 species/stocks are considered overfished, and 5 species/stocks are considered unknown.

```{r plot_hms_stock_statusMAB}
# Plot indicator
ggplotObject <- ecodata::plot_hms_stock_status(report='MidAtlantic')
ggplotObject
```


## Indicator statistics 
Spatial scale: Species-specific; depends on the extent of the managed stock. Can be regional (e.g., Gulf of Mexico or Atlantic blacktip shark) or could be basin-wide (e.g., North Atlantic swordfish, Western Atlantic bluefin)

Temporal scale: Annual summary of stock status

**Synthesis Theme**:

- [X] Regime Shifts


```{r autostats_hms_stock_status}
# Either from Contributor or ecodata
```

## Implications
TBD – our assessment team has asked that the Kobe plots be redone, so please do not include them here yet.

## Get the data

**Point of contact**: [Jennifer Cudney (jennifer.cudney@noaa.gov)](mailto:Jennifer Cudney (jennifer.cudney@noaa.gov)){.email}

**ecodata name**: `ecodata::hms_stock_status`

**Variable definitions**

Current Relative Biomass level =  finalized biomass estimate provided in latest stock assessments Bmsy = biomass at maximum sustainable yield 
Threshold = the threshold identified in respective assessments to indicate overfished vs. overfishing 
International Stock Status, Domestic Stock Status, Years to Rebuild = estimate if provided in assessments Rebuilding Start Date (End Date) = estimate if provided in assessments. 
Current Relative Fishing Mortality Rate = finalized mortality rates provided in latest stock assessment Maximum Fishing Mortality Threshold = reference point identified in latest stock assessment 
International Stock Status, Domestic Stock Status. Stock = the name of the stock assessed Last Assessment Year = the year in which the latest assessment was finalized 
Upcoming Assessment = Year that the next assessment is anticipated (if scheduled) Notes = additional information. Shark Stock = the name of the stock assessed 
Last Assessment Year = the year the latest assessment was finalized 
Last Assessment Type = unless otherwise indicated, this identifies the type of SEDAR assessment used (typically “Benchmark”, “Standard”, or “Update”). This is subject to change as SEDAR adjusts their assessment types. 
Upcoming Assessment = year of upcoming assessment, if scheduled. Upcoming Assessment Type = unless otherwise indicated this identifies the type of SEDAR assessment used. 
Notes = Relevant information concerning the domestic assessments.

```{r vars_hms_stock_status}
# Pull all var names
vars <- ecodata::hms_stock_status |>
   dplyr::select(Var) |>
   dplyr::distinct()

DT::datatable(vars)
```
**Indicator Category**:

- [X] Syntheses of published information


## Public Availability

Source data are publicly available.

## Accessibility and Constraints

Information is publicly available in the most recent Atlantic HMS Stock Assessment and Fishery Evaluation (SAFE) report. Found [here](https://www.fisheries.noaa.gov/atlantic-highly-migratory-species/atlantic-highly-migratory-species-stock-assessment-and-fisheries-evaluation-reports).

**tech-doc link**
<https://noaa-edab.github.io/tech-doc/hms_stock_status.html>

