# Zooplankton Indices {#New Indicator}

**Description**: Model-based abundance indices for zooplankton groups sampled by NEFSC surveys

**Indicator family**: 

- [X] Lower trophic levels


**Contributor(s)**: Sarah Gaichas and Harvey Walsh

**Affiliations**: NEFSC

```{r echo=FALSE}
knitr::opts_chunk$set(echo = F)
library(ecodata)
```
## Introduction to Indicator
Zooplankton are primary consumers in marine ecosystems that transfer energy from phytoplankton to fish, marine mammals, and birds. The Northeast Fisheries Science Center has conducted [zooplankton surveys](https://www.fisheries.noaa.gov/inport/item/9286) since the 1970s. Spatially explicit indices of abundance for several zooplankton groups were estimated using spatio-temporal modeling. 

The groups include: 

**Copepods**

* _Calanus finmarchicus_, = Large copeopds SOE (used in small-large index)

* Large copepods ALL: _Calanus finmarchicus_, _Metridia lucens_, _Calanus minor_, _Eucalanus_ spp., _Calanus_ spp.

* Small copepods ALL: _Centropages typicus_, _Pseudocalanus_ spp., T_emora longicornis_, _Centropages hamatus_, _Paracalanus parvus_, _Acartia_ spp., _Clausocalanus arcuicornis_, _Acartia longiremis_, _Clausocalanus furcatus_, _Temora stylifera_, _Temora_ spp., _Tortanus discaudatus_, _Paracalanus_ spp.

* Small copeopods SOE (used in small-large index): _Centropages typicus_, _Pseudocalanus_ spp., _Temora longicornis_, _Centropages hamatus_

**Euphausiids** 

**Zooplankton volume**

## Key Results and Visualizations
Large and small copepod groups are mainly fluctuating without trend in both spring and fall across the EPUs. However, a significant decrease in small copeopods was apparent in the MAB during fall. Similarly, zooplankton volume is fluctuating without long term trends over time.

Euphausiids, in contrast, show increasing trends over time in all regions during fall, and in the MAB during spring.

**ADD ANY CHANGED RESULTS WHEN SHORT TERM TREND PLOTS ARE IMPLEMENTED**


## Indicator statistics 
Spatial scale: by EPU and full shelf

Temporal scale: Spring (January-June), Fall (July-December)

**Synthesis Theme**:

- [X] Multiple System Drivers


```{r autostats_New Indicator}
# Either from Contributor or ecodata
```

## Implications
Fluctuations in zooplankton community composition can affect both the consumption of phytoplanlton vs the export of energy to the benthos, as well as the availability of prey for zooplankton feeding forage fish, marine mammals, and seabirds. 

The relative dominance of large and small copepods in Northeast US ecosystems has been liked to changes in fish productivity  [@perretti_regime_2017; @Morse2017].

## Get the data

**Point of contact**: [Sarah Gaichas, Sarah.Gaichas@noaa.gov](mailto:Sarah Gaichas, Sarah.Gaichas@noaa.gov){.email}

**ecodata name**: No dataset

**Variable definitions**

Naming key * *Calanus finmarchicus* (calfin_100m3) = “calfin” , * Large copepods (calfin_100m3, mlucens_100m3, calminor_100m3, euc_100m3, calspp_100m3) = “lgcopeALL”, 
* Small copepods (all) (ctyp_100m3, pseudo_100m3, tlong_100m3, cham_100m3, para_100m3, acarspp_100m3, clauso, acarlong_100m3, fur_100m3, ost_100m3, temspp_100m3, tort_100m3, paraspp_100m3) = “smallcopeALL” and 
* Small copepods (SOE) (ctyp_100m3, pseudo_100m3, tlong_100m3, cham_100m3) = “smallcopeSOE”.   Variables **TO BE ADDED ONCE DATA IN ECODATA, NOT ALL DATASETS COULD BE READ INTO FORM**


No Data

**Indicator Category**:

- [X] Extensive analysis, not yet published


## Public Availability

Source data are NOT publicly available.

## Accessibility and Constraints

Request from Harvey Walsh, harvey.walsh@noaa.gov

